---
title: "PA1_template.Rmd"
author: "Vu Tuan Manh"
date: "7/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## I. Loading and preprocessing the data

### 1.1. Load the data

```{r LoadData, message=FALSE}
# Load data in to data frame
activityData = read.csv("Data/activity.csv", header = TRUE)
# Add libraries
library(ggplot2)
library(dplyr)
```

## II. Calculation and plot

<!-- Ignore the missing values in the dataset. -->

### 2.1 Calculate the total number of steps taken per day

```{r activityData, message=FALSE}
stepPerDay <- aggregate(activityData$steps, list(activityData$date), FUN=sum)
colnames(stepPerDay) <- c("Date", "Steps")
```

Information about data

```{r, message=FALSE}
str(stepPerDay)
head(stepPerDay)
```

### 2.2 Make a histogram of the total number of steps taken each day
```{r, message=FALSE, warning=FALSE}
g <- ggplot(stepPerDay, mapping=aes(x=Steps,label=..count..))
g+geom_histogram(boundary=0, binwidth=2500, col="black", fill="lightblue")+
    ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency by day")
```

### 2.3 Calculate and report the mean and median of the total number of steps taken per day
```{r, message=FALSE, warning=FALSE}
# Mean
mean(stepPerDay$Steps, na.rm=TRUE)
```

```{r, message=FALSE, warning=FALSE}
# Median
median(stepPerDay$Steps, na.rm=TRUE)
```

## III. What is the average daily activity pattern?
### 3.1 Make a time series plot (i.e.type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, message=FALSE, warning=FALSE}
# Calculate average steps by interval
groupData <- group_by(activityData,interval)
StepsPerInterval <- summarize(groupData, Steps = mean(steps, na.rm = TRUE))
colnames(StepsPerInterval) <- c("interval", "steps")

# Another option for calculate steps per interval
# StepsPerInterval <- aggregate(steps~interval,data=activityData,FUN=mean,na.action=na.omit)

# draw the line plot
g <- ggplot(StepsPerInterval, aes(interval, steps))
g + geom_line(col="brown")+ggtitle("Average steps per time interval")+
    xlab("Interval")+ylab("Steps")
```

### 3.2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, message=FALSE, warning=FALSE}
subset(StepsPerInterval, steps==max(StepsPerInterval$steps))
```

## IV. Imputing missing values
### 4.1 Calculate and report the total number of missing values in the dataset
```{r, message=FALSE, warning=FALSE}
# Number of rows with N/A value 
nrow(subset(activityData,is.na(activityData)))
```

### 4.2 Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
### We will fill missing value by using the mean steps in each interval
### 4.3 Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r, message=FALSE, warning=FALSE}
# clone data frame
activityDataComplete <- activityData
# Matching and update missing steps with mean value of appropriate interval
activityDataComplete$steps <- ifelse(is.na(activityDataComplete$steps),
    round(StepsPerInterval$steps[match(activityDataComplete$interval,
    StepsPerInterval$interval)],0), activityDataComplete$steps)
```

### 4.4 Make a histogram of the total number of steps taken each day and Calculate 
```{r, message=FALSE, warning=FALSE}
# Calculate new mean step per day
stepPerDayNew <- aggregate(activityDataComplete$steps, list(activityDataComplete$date), FUN=sum)
colnames(stepPerDayNew) <- c("Date", "Steps")
# draw the line histogram for new step per day data (after fill missing value)
g <- ggplot(stepPerDayNew, mapping=aes(x=Steps,label=..count..))
g+geom_histogram(boundary=0, binwidth=2500, col="black", fill="lightblue")+
    ggtitle("Histogram of steps per day (no missing value)")+xlab("Steps")+ylab("Frequency by day")
```

### And report the Mean and Median total number of steps taken per day
```{r, message=FALSE, warning=FALSE}
# Calculate and report the mean and median total number of steps taken per day
# Mean steps per day (no missing value)
mean(stepPerDayNew$Steps)
# Mean steps per day (with missing value)
mean(stepPerDay$Steps, na.rm=TRUE)
# Median steps per day (no missing value)
median(stepPerDayNew$Steps)
# Mean steps per day (with missing value)
median(stepPerDay$Steps, na.rm=TRUE)
```

### Compare total steps per day after imputing missing value
```{r, message=FALSE, warning=FALSE}
totalstepPerDay <- sum(stepPerDay$Steps, na.rm=TRUE)
totalstepPerDayNew <- sum(stepPerDayNew$Steps)
paste("Total steps (with missing value): ",totalstepPerDay," steps",sep="")
paste("Total steps (no missing value): ",totalstepPerDayNew," steps",sep="")
paste("Deviation of steps (after - before imputing: ", totalstepPerDayNew- totalstepPerDay, sep="")
```

## V.Are there differences in activity patterns between weekdays and weekends?
### 5.1 Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r, message=FALSE, warning=FALSE}
# Convert to date format
activityData$date <- as.Date(activityData$date, format= "%Y-%M-%d")
# Create weekday variable
activityData$weekday <- weekdays(activityData$date)
# Create day type for weekday and weekend
activityData$dayType <- ifelse(activityData$weekday=='Saturday' | 
    activityData$weekday=='Sunday', 'weekend','weekday')
# preview head data
head(activityData)
```

### 5.2 Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r, message=FALSE, warning=FALSE}
# Create data frame with steps per time interval for weekday and weekend
stepPerTimeDayTye <- aggregate(steps~interval+dayType,data=activityData,FUN=mean,na.action=na.omit)
# draw the line plot
g <- ggplot(stepPerTimeDayTye, aes(interval, steps))
g + geom_line(col="blue")+ggtitle("Average steps per time interval: weekdays vs. weekends")+
    xlab("Interval")+ylab("Steps")+
    theme(plot.title = element_text(face="bold", size=12))+
    facet_grid(dayType ~ .)
```







